---
title: "Motor Trend Analysis Report"
author: "Arman Iskaliyev"
date: "10/03/2018"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(GGally)
library(datasets)
set.seed(198712)
```

##1. Executive Summary
This report was developed for the Motor Trend Magazine to examine which type of automobile transmission has better result in mileage (miles per gallon or mpg). The report provides an analysis and answers for exploring questions. Methods of analysis include fitting regression models by using stepwise regression method, using T-test for hypothesis test. All R calculations and exploratory plots used in this report can be found in the appendices part. Results of analysis show that manual transmission type is better for mileage than automatic transmission type, and difference between automatic and manual transmissions is 1.8 MPG.

##2. Questions of interest

- **Is an automatic or manual transmission better for MPG**

- **Quantify the MPG difference between automatic and manual transmissions**

##3. Data Processing
###3.1 Dataset description
The `mtcars` dataset from base R was used for this analysis. The data in dataset was extracted from the 1974 Motor Trend US magazine, and comprises fuel consumption and 10 aspects of automobile design and performance for 32 automobiles (1973–74 models).

Variable  | Desc
----------|--------------------------
mpg	     |Miles/(US) gallon
cyl	     |Number of cylinders
disp	     |Displacement (cu.in.)
hp	     |Gross horsepower
drat	     |Rear axle ratio
wt	     |Weight (1000 lbs)
qsec	     |1/4 mile time
vs	     |V/S
am	     |Transmission (0 = automatic, 1 = manual)
gear	     |Number of forward gears
carb	     |Number of carburetors

###3.2 Loading and exploring dataset
```{r loaddata}
data("mtcars")
str(mtcars)
summary(mtcars)
```

###3.3 Transformations
Some variables look like a categorical but they was saved as numeric. We converted them into factors manually.   
```{r asfact}
vars_to_fctr <- c("am","cyl","carb","vs","gear")
#ds[vars_to_fctr] <- lapply(ds[vars_to_fctr], factor)
ds <- mtcars %>% 
     tibble::rownames_to_column() %>%
     mutate_at(vars_to_fctr, funs(factor(.)))

levels(ds$am) <- c("automatic", "manual")
levels(ds$vs) <- c("V", "S")
str(ds)
```

##4. Exploratory Data Analysis
First we calculated correlation coefficients for our outcome variable, a `mpg`, vs. all other numeric variables. 
```{r cormatrix}
ggcorr.out <- ds %>% select_if(is.numeric) %>%
     ggcorr(label = TRUE)
```

We found weak correlations in `qsec` and `drat` variables. So these variables will not be needed us in further analysis and they could be excluded from dataset.
```{r corrview}
ggcorr.out$data %>% filter(y == "mpg") %>% arrange(abs(coefficient)) %>% select(-label)
```

We also build boxplot pairs of `mpg` vs. all categorical variables. Here we saw robust relationship between `mpg ~ cyl`, `mpg ~ am` and `mpg ~ vs`, but there are some outliers here, in 8-cylinder observation (Cadillac Fleetwood, Lincoln Continental, Pontiac Firebird) and V type engine observation (Porsche 914-2). We decided to exclude `carb` and `gear` variables.
```{r outliers}
var_to_drop <- c("qsec", "drat", "gear", "carb")
ds %>% filter((vs == "V" & mpg > 25) | ((cyl == "8" & mpg < 12) | (cyl == "8" & mpg > 19)))
ds <- ds %>% select(-one_of(var_to_drop), -rowname)
```

##5. Question 1. Is an automatic or manual transmission better for MPG
As we just have seen in exploratory plot, there is strong relationship between mileage and transmission type, and manual transmission type looks better for MPG. To prove our assumption we conducted two-sided T-test.
```{r ttest}
(ttest.out <- t.test(mpg~am, data = ds))
```
###Interpretation
So calculated p-value `r ttest.out$p.value` allow as to reject null hypothesis, and prove that manual transmission is better for MPG.

##6. Question 2. Quantify the MPG difference between automatic and manual transmissions
###6.1. Fit multiple models and Best model selection
Here we use stepwise regression method to fit multiple models and choose best model. We started with multivariate model including all variables.

Fit initial model:
```{r model_init}
mod.init <- lm(mpg ~ ., data = ds)
summary(mod.init)
```

Apply step function with backward direction to exclude variables one at time from formula, and to fit them. At the end of the function it give us best fitted model:  
```{r model_best}
mod.best <- step(mod.init, direction = "backward", trace = 0)
summary(mod.best)
```
###Interpretation
According to the results of the stepwise method the best regression model for mpg includes all these variables:

* cyl
* hp
* wt
* am

This model has a quite high R-squared value equal to `r summary(mod.best)$r.squared`. Our model show that manual transmission type is 1.80 better than automatic.

##7. Conclusion
* Manual transmission type is better for mileage than automatic
* Manual transmission type is better for mileage by a factor of 1.8 than automatic

##Appendices
### Exploratory plots
Relationship between mileage and transmission type.
```{r boxplot}
qplot(x = am, y = mpg, data = ds, geom = "boxplot")
```

Correlation of mpg vs. all numerical variables. 
```{r cormatrixplot}
ggcorr.out + ggtitle("Correlation matrix")
```

Matrix of scatterplots mpg vs. all categorical variables.
```{r cormatrixsctplot}
ds %>% select_if(is.factor) %>%
     bind_cols(select(ds, mpg)) %>% 
     select(mpg, everything()) %>%
     ggpairs()
```

Residual plots of best model
```{r residplot}
par(mfrow=c(2,2))
plot(mod.best)
```

